### 2014-01-07 msc-thesis/chapters/4/chapter-discussion.tex
MSC base discussions now in place.

@@

@@

As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:review}. This inability to generalize is what we believe to be the
leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
information being presented to them, rendering the system useless. These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's 
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

@@

Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
in that they result in a trade off 
of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.

@@

%mature point, static analysis
\textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
some software project configurations and software languages to be quite prone to indirect conflicts.}

@@

A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
research than may have been previously thought. Static analysis may be useful in regards to indirect conflict context in that we
could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
changes occur at a given point in time which could be used for determining such a context of mature points.

This added level of context, through the notion of a mature point, 
would add to the prevention versus cure debate as previously discussed. Prevention may be a better
choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
causes more bugs than can be prevented. These possibilities existing, it may be imperative to discover more of project stability
and the mature point in order to fully understand the nature of indirect conflicts and their context.

@@

Aside from context, dependency identification and tracking is a key missing component of
indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and 
organizational structures of software teams. Between increased modularization (multiple sub projects or repositories), 
cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have 
become quite limited in identifying and tracking dependencies where they were once strong.

Relational database schemas are one of the highest sources of indirect
conflicts found in their projects, ``it breaks stuff all over the place'',
yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
been scarcely researched in terms of indirect conflicts. This falls in with our previous understanding of cross language
dependency tracking in that database schemas are independent of languages which may be on the receiving end of their output.

>>>>

\textbf{The Human Factor of Indirect Conflicts} \textit{An important component of indirect conflicts are the developers
themselves and how their notions of other's work is perceived across a project.}

As we have seen through Study 1, developers which are tied to source code objects can become a focal point of
indirect conflicts through the life of a project. For instance we can see from Table~\ref{tab:ratio} that
developers Daniel and Anne on Hibernate-ORM are always linked indirect conflicts when dependencies between
their source code objects change. This is an important observation to make when moving forward with indirect
conflicts.

What is really happening between these two developers is the real issue to consider. Daniel could have an
assumption about the way Anne's code works which causes Anne's changes to have negative impacts on Daniel's
own work. For instance, Daniel may expect a method of Anne's to have a special return case, which may be correct,
however when Anne changes that special case or removes it, Daniel's code can be negatively affected. In this extreme
result found in Study 1, further analysis showed that one central method in the software project was being changed
frequently and causing Daniel's code to be negatively affected in some way.

The question of how to prevent this type of negative impact is ultimately the goal of indirect conflicts. Study 1 has
shown that pairs of developers can often be a large component to that goal as well. (Impact from Study 2 was
created the address just this issue.)

\textbf{Information Overload} \textit{My tool Impact, as well as many other indirect conflict tools, suffer from
information overload in their delivery to developers which is a key issue in preventing adopting and acceptance
of indirect conflict tools.}

As was previously stated, many indirect conflict tools end up suffering from information overload to developers
and other end users. Impact was initially created to take the knowledge learned from Study 1 and attempt to create
a new indirect conflict system based on developer interactions inside the code base which would potentially address
information overload. However, we now know from the results of Study 2, that Impact once again suffered defeat to information
overload based on the case study evaluation. Ultimately, this sense of information overload ends up causing
adoption of indirect conflict tools to fail which in turn causes some research components to have failed as
well.

From previous works, as well as from Study 2, we know that information overload is a large issue in indirect
conflicts. It has been found that a large number of dependencies in software caused by the nature large software
projects is a root issue in information overload~\cite{Sarma:2007:TSA,Servant:2010:CPI}. These large sizes in
dependencies are ultimately what cause information overload as dependencies cannot be evaluated as to their
relevance in a certain source code change with ease. In other words, we cannot determine which of the numerous
dependencies will fail (outside of compilation and testing failures) when source code is changed.

A potential solution derived from the evaluation of Impact comes from the ideas of Meyer~\cite{Meyer:1988} on 
``design by contract''. In this methodology, changes to method preconditions and postconditions are considered 
to be the most harmful. This includes adding conditions that must be met by both input and output parameters 
such as limitations to input and expected output. To achieve this level of source code analysis, the ideas of 
Fluri et al.~\cite{Fluri:2007:CDT} can be used on the previously generated ASTs for high granularity source code 
change extraction when determining if preconditions or postconditions have changed. While this solution does not 
wholly address the problems of information overload as previously stated through failures in dependencies, it does 
reduce the number of source code changes to  be analyzed which in turn will reduce the amount of information on 
the whole which is put forth by the indirect
conflict system. This solution however does also run the risk of missing more indirect conflicts as many conflicts
can occur outside of changes to method contracts. This solution represents my first ideas of fixing information overload in regards to
indirect conflicts and is again found in the later discussion of root causes of indirect conflicts found in
the next section.

@@

While the previous section discussed the main motivations for taking a step back on indirect conflicts in order to understand
better what can be done to improve developer work flow, this section will discuss exactly those steps back. From Studies 3 and
4 we will notice 3 main trends that have been discovered for indirect conflicts in both industry and in academia. These 3 major
trends to now be discussed are: unwanted awareness, prevention versus cure, and the gap in software evolution analysis. This
section will also include a brief discussion regarding the implications, learned from studies 3 and 4, for both academia and
for industry through tools.

@@

As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:review}. This inability to generalize is what I believe to be the
leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
information being presented to them, rendering the system useless. These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's 
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

@@

Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
to measure. The pros and cons of prevention versus cure I have identified, are similar to those of unwanted awareness,
in that they result in a trade off 
of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.

@@

%mature point, static analysis
\textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
and analysis unfriendly languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
some software project configurations and software languages to be quite prone to indirect conflicts.}

@@

A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
research than may have been previously thought. Static analysis may be useful in regards to indirect conflict context in that we
could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
These change trends are exactly what Study 4 set out to address.

Through Study 4, I showed that 9 critical change trends exist at a major release of a variety of open source projects. This
knowledge could help identify, as was previously stated, when indirect conflicts are more likely to occur in a project or not.
For example, the trend of adding test methods was found to occur more after a major release than before. This trend along with
the knowledge that indirect conflicts occur more at the start of a new development cycle can be used to predict times in development
where indirect conflicts are likely to occur automatically. An automated system may see that more test methods are being added than
before in the past 4 weeks or so and be able to alert developers to an instability in the code which in turn means an elevated
risk of indirect conflicts.

Similarly, the trend of adding private classes is seen to occur more so before a major release than after. This knowledge again
can be used in an automated system along with the fact that indirect conflicts are less likely to occur towards the end of a development
cycle. The automated system will be able to identify a stability in the project through private classes being added and alert developers
that the code is stabilizing and that large changes to the project should be avoided until the start of the next development cycle
in order to further reduce the risk of indirect conflicts.

This added level of context, through the notion of a mature point, 
would add to the prevention versus cure debate as previously discussed as well. Prevention may be a better
choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
causes more bugs than can be prevented. These possibilities existing, it may be imperative to discover more of project stability
and the mature point in order to fully understand the nature of indirect conflicts and their context.

@@

In regards to these contextual identifications in software projects, dependency identification and tracking is a key missing component of
indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and 
organizational structures of software teams. Between increased modularization (multiple sub projects or repositories), 
cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have 
become quite limited in identifying and tracking dependencies where they were once strong. As software has become more
sophisticated over time, static analysis tools have gone from extremely useful to only occasionally useful. Since many
projects involve several languages, sub projects, and database schemas, static analysis has become cripplingly obsolete
in the industry of today. In order to move indirect conflicts and many other research areas forward, this gap of cross domain
static analysis must be filled.

As an example, relational database schemas are one of the highest sources of indirect
conflicts found in their projects, ``it breaks stuff all over the place'',
yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
been scarcely researched in terms of indirect conflicts. This falls in with my previous understanding of cross language
dependency tracking in that database schemas are independent of languages which may be on the receiving end of their output.
